---
title: "R ggplot2 Assignment 2"
author: Yan Li
date: Aug 8th 2015
output: word_document
---

We first have to load the Required packages.

```{r}
#Load required packages in the user-defined library location
.libPaths("C:/work/Rpackages/R3.2.1")
suppressMessages(library(ggplot2, quietly = T))
suppressMessages(library(dplyr, quietly = T))
suppressMessages(library(mosaicData, quietly = T))

#turn off warning messages
options(warn=-1)
```


The data used for this assignment is mpg or diamonds from ggplot2 package.

```{r}
#Show the top 6 rows of the mpg/diamonds data to know the data information
head(mpg)
head(diamonds)
```

This assignment is to create 3 plots, each to answer one question and with 3-5 iterations to show the process.

The first question is about the diamonds data: what is the price change along with the carat for different cut and clarity.

First we see the price change with shape = cut and color = clarity:

```{r ggplot1, fig.width = 8, fig.height = 6, message=FALSE }
#Basic Model
p <- ggplot( data=diamonds, aes(x=carat, y=price, shape = cut, color= clarity) ) 
#Scatter plot 
 p+ geom_point( alpha=0.1 )
```

It is not very clear to see the price change if the customer has a clear requirement of clarity and cut. And it takes a long time to get the plots (over 5 mins). So we can subset the data into different group based on clarity and cut:


```{r ggplot2, fig.width = 8, fig.height = 6, message=FALSE }
#Delete the shape to speed up the generation of the plot 
p2 <- ggplot( data=diamonds, aes(x=carat, y=price, color= color ) )

#Subset data to group based on clarity and cut
#First we check facet = cut~.
p2+ geom_point( alpha=0.3, size = I(2))+
   facet_grid(cut~. )
#Second we check facet = clarity ~.
p2+ geom_point( alpha=0.3, size = I(2))+
   facet_grid(clarity~. )
#Last facet = clarity ~cut
p2+ geom_point( alpha=0.3, size = I(2))+
   facet_wrap(clarity~cut )

#Add 2d density
#Add title
 p2+ geom_point( alpha=0.3, size = I(2))+
   facet_wrap(clarity~cut )+stat_density2d(color = "black")  +
  labs(title="Price vs. Carat", x="Carat of diamond", y="Price of Diamonds")
```

From the above plots, customers can easily find any combination of clarity and cut subgroup. Then they can further examine the price change.

Second question is: How many choices do I have if my bugget of buying a diamond is $10,000-$12,000 (the range of the diamonds data  for price is $326-$18,823)? 

This needs to limit the price range.

```{r ggplot3, fig.width = 8, fig.height = 6, message=FALSE }
##limit data range and set 'outlier' color
require(scales)
p3 <- ggplot( data=diamonds, aes(price, carat) ) 

p3+
geom_point(aes(color=price), size=2.5,alpha = 0.3) +
   scale_color_gradient2(low=muted("light blue"), high=muted("dark blue"), mid="blue",
                        midpoint = 11500, limits=c(10000,12000), na.value="gray")
```


Now We focus on the the price range(10000,12000).

```{r ggplot4, fig.width = 8, fig.height = 6, message=FALSE }

p3+geom_point(aes(colour = color))+
  scale_x_continuous(limits=c(10000,12000))

#Add shape, size and alpha to make a better looking plot
p3+geom_point(aes(colour = color, shape= cut, size = 2.5, alpha = 0.3))+
    scale_x_continuous(limits=c(10000,12000))

```

It is not easy to find cut,clarity features. So subsetting the data by clarity. 

```{r ggplot5, fig.width = 8, fig.height = 6, message=FALSE }
  p3+geom_point(aes(colour = color, shape = cut,size = 2.5, alpha = 0.3))+
  facet_grid(clarity~.)+
   scale_x_continuous(limits=c(10000,12000))+
 labs(title="Daimonds Price $10000-$12000", x="Price of Diamonds", y="Carat of diamond")
```

Now we may want to limit the weight of the diamonds too:

```{r ggplot6, fig.width = 8, fig.height = 6, message=FALSE }
#Limit the carat
p3+geom_point(aes(colour = color, shape = cut,size = 2.5, alpha = 0.3))+
  facet_grid(clarity~.)+
   scale_y_continuous(limits=c(0.5,1.5))+
  scale_x_continuous(limits=c(10000,12000))+
  labs(title="Daimonds Price $10000-$12000; Weight 0.75-1.5 carat", x="Price of Diamonds", y="Carat of diamond")
```

Now we can choose the diamonds based on our buget.

The third plot is about: How does the geometric measures(table) of diamonds relate to the weight (carat)

```{r ggplot7, fig.width = 8, fig.height = 6, message=FALSE }
#2d histogram using hexagonal shape
require(hexbin)
p4 <- ggplot( diamonds, aes(x = carat, y=table) )

p4 + stat_binhex()

#2d kernel density
p4 + stat_density2d()

#combine the above two
p4 + 
  stat_binhex(aes(fill=..density..)) +
  stat_density2d(color="gray75", size=.3)

#Use tile showing the density
p4  +xlim(1,3)+
 # stat_binhex(aes(fill=..density..)) +
  stat_density2d(geom = "tile",aes(fill=..density..),contour=F)+
scale_fill_gradient(limit = c(1e-5,0.45))

```

