---
title: "Assignment 4"
author: Yan Li
date: August 21 2015
output: word_document
---

#Load packages, data and set the working library:

```{r}
.libPaths("C:/work/Rpackages/R3.2.1")
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(mosaicData))
suppressMessages(library(babynames))
suppressMessages(library(fastR))
suppressMessages(library(tidyr))
suppressMessages(library(ggvis))
suppressMessages(library(scales))

#turn off warning messages
options(warn=-1)

#Set working directory
setwd("C:\\work\\Rggplot2\\Lesson 4")
```

The data used in this assignmet is Texas:

```{r}
Texas <- read.csv("texas.csv")
head(Texas,1)

#Simplified the data: delete duplicated rows
PopChoro <- unique(Texas[c("county", "long", "lat", "group", "order", "population", 
"bin")])
head(PopChoro,2)

```


The original plot of Texas data showing the population by county:

```{r qplot1, fig.width = 10, fig.height = 8, message=FALSE}
q <- qplot(long, lat, data = PopChoro, geom = "polygon", group = group, fill = bin)

q
```

This assignment is to improve the plot by:

1 Adding an informative title

2 Choosing a better color scale

3 Giving more informative title and keys on     the legend

4 Removing the unnecessary x and y axis labels

5 Making it look more like a map by giving it a white background.

6 (Extra credit) Use a better coordination system (skim Chapter 6 "Positioning" for some ideas)

#1 Adding an informative title

```{r qplot2, fig.width = 10, fig.height = 8, message=FALSE}
#Add texts by labs()
myLabs<- labs(title="Population of Texas by County",
x = "longitude of Texas (long)",
y = "lattitude of Texas (lat)",
fill = "population range (bin)")

q+myLabs

#Add some fomata and colors
myTheme<- theme(plot.title = element_text(colour='navy', size=18, face='bold', family='serif'),
  axis.title.x = element_text(colour='blue', size=14,  family='serif'),
         axis.title.y = element_text(colour='blue', size=14,  family='serif')
  )

q+myLabs+myTheme
```

#2 Choosing a better color scale

```{r qplot3, fig.width = 10, fig.height = 8, message=FALSE}
#Use qulitative red pallete to show the populations change

# qplot(long, lat, data = PopChoro)+ geom_polygon(aes(group= group, fill = bin))+scale_fill_brewer(type = "qual")

myScale <- scale_fill_brewer(type = "qual", palette = "Reds")

q+myLabs+myTheme+myScale
```

#3 Giving more informative title and keys on the legend

The title was modified in question 1 by labs(). The following change the font color and size.

```{r qplot4, fig.width = 10, fig.height = 8, message=FALSE}

myTheme2 <- theme(legend.title = element_text(colour='red', size=10,  family='serif'),
                  legend.text = element_text(colour='red', size=10,  family='serif'))

q+myLabs+myTheme+myScale+myTheme2
```

#4 Removing the unnecessary x and y axis labels

```{r qplot5, fig.width = 10, fig.height = 8, message=FALSE}

myTheme3 <- theme(axis.text.x = element_blank(),
axis.ticks.x = element_blank(),
axis.text.y = element_blank(),
axis.ticks.y = element_blank())

q+myLabs+myTheme+myScale+myTheme2+myTheme3
```

#5 Making it look more like a map by giving it a white background.

```{r qplot6, fig.width = 10, fig.height = 8, message=FALSE}
myTheme4 <- theme(panel.grid.major=element_line(colour="white"),
panel.grid.minor=element_line(colour="white"),
panel.background=element_rect(fill=alpha("white",0.2)))

q+myLabs+myTheme+myScale+myTheme2+myTheme3+myTheme4
```
